The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-16 paper/ref_approach.tex
minor edits + added references

In statistical applications, one of the main steps in the modeling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals. 
First, if the covariates themselves are of interest,
we can use variable selection as an inferential tool to find the entire 
collection of variables which have some predictive information about the target variable.
Second, as simpler models come with the advantages of reduced
measurement costs and improved interpretability, we may be interested in
finding the minimal subset of covariates which
provides good predictive performance (or good balance between simplicity and
predictive performance). 
In the latter case, where the focus is on prediction, the true data generation
mechanism of future data can be approximated either by using the observed data directly or
alternatively by using predictions from a \emph{reference model}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression, the observed empirical data distribution is utilized as a
proxy of future data usually in combination with cross-validation
or information criteria to provide estimates for out-of-sample
predictive performance.
In contrast, reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, a full-encompassing model including all
covariates. The reference
model approach has been used for long time in Bayesian statistics
since the seminal work of \citet{paper:reference_lindley}. For more
historical references, see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.
Reference models have been also used in non-Bayesian contexts,
where \cite{harrell2015regression} describes them as full models that
can be thought of as a ``gold standard'' (for a given application).
For example,
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. 
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed responses $y$.

In this paper, we examine the benefits of using reference
models in detail. We illustrate how a properly designed reference model is able 
to filter parts of the noise present in the data, and hence provides an improved and more
stable selection process. This holds even if the reference model does not actually resemble the
true data generating process. Moreover, our analyses indicate that the substantial reduction of
variance attributable to noise is usually
more important than small potential bias due model misspecification.
We argue that, regardless of how the reference model is set up and used in the
inference prodecure, it can be always seen as acting as a filter on the observed 
data. Following this idea, we devise a simple reference model approach that can
be applied on top of any variable selection procedure.

In our simulations and analysis of real data, we apply several state-of-the-art methods for variable selection and compare,
for each of these methods, the results obtained by using a data based and a
reference model based approach. When a full 
variable selection is carried out, we report and compare quantities such as 
the false discovery rate, sensitivity and stability of the selection. For fully Bayesian methods, we additionally report results in terms of shrinkage and accuracy of the posterior estimates.
We show the benefits
of a reference model approach, regardless of what specific procedure is applied, using 
the normal means estimation problem as a benchmark. 
Variable selection is typically most challenging in 
\textit{small n large p} scenarios, that is, scenarios with small number of
observations but a lot of unknown model parameters. These occur frequently, for example, in microarray data analysis in medicine or biology.
In general, the less data we have and the more complex 
the estimated models, the higher the benefit of using a reference
models as the basis for variable selection.
Our results indicate that the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures. We want to note that, although we focus on
Bayesian approach in this paper, the benefits of the reference model 
approach can also be obtained in non-Bayesian settings.

>>>>

In statistical applications, one of the main steps in the modeling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals. 
First, if the covariates themselves are of interest,
we can use variable selection as an inferential tool to find the entire 
collection of variables which have some predictive information about the target variable.
Second, as simpler models come with the advantages of reduced
measurement costs and improved interpretability, we may be interested in
finding the minimal subset of covariates which
provides good predictive performance (or good balance between simplicity and
predictive performance). 
Whenever the predictive capability is guiding the selection, the true data generation
mechanism of future data can be approximated either by using the observed data directly or
alternatively by using predictions from a \emph{reference model}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward 
regression \citep{venables2013modern,harrell2015regression}, the observed empirical data distribution is utilized as a
proxy of future data usually in combination with cross-validation
or information criteria to provide estimates for out-of-sample
predictive performance.
In contrast, reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, a full-encompassing model including all
covariates. The reference
model approach has been used for long time in Bayesian statistics
since the seminal work of \citet{paper:reference_lindley}. For more
historical references, see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.
Reference models have been also used in non-Bayesian contexts,
where \cite{harrell2015regression} describes them as full models that
can be thought of as a ``gold standard'' (for a given application).
For example,
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. 
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed responses $y$.

In this paper, we examine the benefits of using reference
models in detail. We illustrate how a properly designed reference model is able 
to filter parts of the noise present in the data, and hence provides an improved and more
stable selection process. This holds even if the reference model does not actually resemble the
true data generating process. Moreover, our analyses indicate that the substantial reduction of
variance attributable to noise is usually
more important than small potential bias due model misspecification.
We argue that, regardless of how the reference model is set up and used in the
inference procedure, it can be always seen as acting as a filter on the observed 
data. Following this idea, we devise a simple reference model approach that can
be applied on top of any variable selection procedure.

In our simulations and analysis of real data, we apply several state-of-the-art methods for variable selection and compare,
for each of these methods, the results obtained by using a data based and a
reference model based approach. When a full 
variable selection is carried out, we report and compare quantities such as 
the false discovery rate \citep{paper:fdr_BH}, sensitivity and stability of the selection \citep{paper:stability}. For fully Bayesian methods, we additionally report results in terms of shrinkage and accuracy of the posterior estimates.
We show the benefits
of a reference model approach, regardless of what specific procedure is applied, using 
the normal means estimation problem as a benchmark. 
Variable selection is typically most challenging in 
\textit{small n large p} scenarios, that is, scenarios with small number of
observations but a lot of unknown model parameters. 
These occur frequently, for example, 
in microarray data analysis in medicine or biology \citep{dudoit2003multiple,liao2007logistic,paper:efron}.
In general, the less data we have and the more complex 
the estimated models, the higher the benefit of using a reference
models as the basis for variable selection.
Our results indicate that the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures. We want to note that, although we focus on
Bayesian approach in this paper, the benefits of the reference model 
approach can also be obtained in non-Bayesian settings.
